### 2004-04-27 Documents/STFormat.tex
Representation of Time revised

%% modified by Thomas Behr
% the numeric type real is added
% restriction of direction for cycles is removed
% the types Date and Datetime are added
% the representation for <map_point_value> is changed (numeric problems)
% 23.4.2004  another representations for instant included

\documentclass[english,a4paper]{article}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{a4}
\usepackage{amsmath}
\headheight-1cm
\headsep0mm
\textheight25.5cm

%\makeatletter

\sloppy

\usepackage{babel}
%\makeatother

\newcommand{\R}{\ensuremath{\mbox{I\hspace*{-0.1em}R}}}
\begin{document}

\title{External Representation of Spatial and Spatio-Temporal Values}

\author{Jose Antonio Cotelo Lema \\{\small extended by Thomas Behr}}

\maketitle
\begin{quote}
%%\begin{center}Last change: 17.03.2000\end{center}
%%\begin{center}Last change: 18.12.2003\end{center}
\begin{center}Last change: 23.4.2004\end{center}
\end{quote}

\section{Introduction}

For supporting the diffusion and interchange of any kind of information
between systems, a well defined and unambiguous specification of an
external representation for such information is required. Therefore,
we need to specify an external representation for the spatial and
spatio-temporal information used by the \emph{STDB} package. The main
goal in such a format must be to allow the exchange of this information
between completely different platforms.

\subsection{Notation}

In the definitions of the external representation of spatial and spatio-temporal
data types in the \emph{STDB} package, the following nomenclature
will be used:

\begin{description}
\item [<x>]a value \emph{x}, represented using the string representation
corresponding to the type of such a value.
\item [<x$_{1}$>~<x$_{2}$>~$\cdots $~<x$_{n}$>]a list of \emph{n}
values. If nothing explicit is said, \emph{n}$\, \geq \, $0. Note
that \emph{n}$\, =\, $0 means an empty list.
\end{description}

\section{Numerical Representation}

For the representation of spatial and spatio-temporal data, a relevant
issue is the representation of their numeric values (used for representing
coordinates, instants in time or functions from the time domain to
the space domain). A requirement is that the representation used for
such numeric values must be able to represent their values exactly,
so the process of exporting and importing data will be safe and will
allow (for example) to restore the data that we have previously exported.

For ensuring the robustness and closure of spatial and spatio-temporal
operations, the use of rational numbers for representing coordinates
and time is required, and the precision used in such numbers could
be arbitrarily large. Therefore, we need to specify how the numeric
values of spatial and spatio-temporal types must be represented in
such cases. For this purpose we will define the external representation
of an auxiliary type \emph{Numeric}, to which belong the coordinates
in space, time instants and the coefficients of spatio-temporal functions.
When representing a spatial or spatio-temporal value, it is allowed
to mix the different representation formats for the \emph{Numeric}
values in it, using for each \emph{Numeric} value the more appropiate
one.

\paragraph{Integer numbers:}

An integer \emph{Numeric} value \emph{I} can be represented in two
different ways, depending on the size of the integer value:

\begin{itemize}
\item \textbf{Signed integers of 32 or less bits:} in this case it can be
represented just as the string representation of the integer number,
optionally with sign (if no sign appears, positive is assumed).

\begin{quote}
\textbf{NOTE:} such a representation of an integer number does NOT
enclose the integer value in brackets.
\end{quote}
\item \textbf{Signed integers of more than 32 bits:} this category includes
arbitrary precision integers. In this case the integer number is represented
as an integer in base $2^{32}$, in the form:

\begin{quotation}
( largeint <sign> <size> <m$_{size-1}$> $\cdots $ <m$_{0}$> )
\end{quotation}
where:

\begin{itemize}
\item <sign> can be either the symbol {}``+'' (positive number), {}``-''
(negative number) or nothing (interpreted as positive).
\item <size> is an integer number (its string representation) defining the
number of elements \emph{m$_{i}$} appearing after it. \emph{<size>}
must be bigger than 0.
\item \emph{<m$_{i}$>} is the string representation of a 32 bits unsigned
integer.

The value \emph{I} represented is:

\begin{description}
\item[] \emph{I$\, =\, $sign\_value$\, *\, $(m$_{size-1}$$*$$\left(2^{32}\right)^{size-1}$$+$$\cdots $$+$}\emph{m$_{1}$$\, *$$\left(2^{32}\right)^{1}$$+\, $m$_{0}$)}
\end{description}
where \emph{<sign\_value>} is 1 if \emph{<sign>} is positive ({}``+''
or nothing) and -1 if \emph{<sign>} is negative ({}``-'').

\end{itemize}
Integer numbers of 32 or less bits can (optionally) be represented
using the format for integers of more than 32 bits.

\end{itemize}

\paragraph{Real numbers:}
\label{real_numbers}
A real number is represented as the string representation of the real number.
The string is formed in the familar way used in the most programming languages
like C or Pascal. Because no cast function exists, a dot or an exponent is required
to distinguish a real number from an integer.

\paragraph{Rational numbers:}

A rational \emph{Numeric} value \emph{R} is represented as:

\begin{quotation}
( rat <sign> <intPart> <numDecimal> / <denomDecimal> )
\end{quotation}
where

\begin{itemize}
\item \emph{<sign>} can be either the symbol {}``+'' (positive number),
{}``-'' (negative number) or nothing (interpreted as positive).
\item \emph{<intPart>}, \emph{<numDecimal>} and \emph{<denomDecimal>} are
non negative ($\geq 0$) integer \emph{Numeric} values (each of them
in any of the formats defined above), \emph{<numDecimal>}$\, <\, $\emph{<denomDecimal>}
and g.c.d.(\emph{<numDecimal>}, \emph{<denomDecimal>})$\, =\, $1
(if \emph{<numDecimal>}$\, =\, $0 then \emph{<denomDecimal>}$\, =\, $1).
\end{itemize}
The value of \emph{R} represented is:

\begin{quotation}
\emph{R}$\, =\, $\emph{sign\_value $*$} (\emph{intPart}$\, +\, \frac{\textrm{numDecimal}}{\textrm{denomDecimal}}$)
\end{quotation}
where \emph{<sign\_value>} is 1 if \emph{<sign>} is positive ({}``+''
or nothing) and -1 if \emph{<sign>} is negative ({}``-'').

If the value of \emph{R} is in fact an integer value, it can (optionally)
be represented as an integer \emph{Numeric} value.

\section{Spatial Values}

\subsection{Point}

\begin{quotation}
(\emph{<xCoord>} \emph{<yCoord>})
\end{quotation}
where \emph{<xCoord>} and \emph{<yCoord>} are the \emph{X} and \emph{Y}
coordinates of the point respectively. Both are \emph{Numeric} values.

\subsection{Points}

\begin{quotation}
((<x$_{1}$> <y$_{1}$>) (<x$_{2}$> <y$_{2}$>) $\cdots $ (<x$_{n}$>
<y$_{n}$>))
\end{quotation}
where (\emph{<x$_{i}$> <y$_{i}$>}) represents a point value and
\emph{n}$\, \geq \, $0.

\subsection{Line}

\begin{quotation}
(<seg$_{1}$> <seg$_{2}$> $\cdots $ <seg$_{n}$>)
\end{quotation}
where \emph{<seg$_{i}$>} is a segment value and \emph{n}$\geq $0.

A segment value is represented as:

\begin{quotation}
(<x$_{1}$> <y$_{1}$> <x$_{2}$> <y$_{2}$>)
\end{quotation}
where \emph{<x$_{i}$>} and \emph{<y$_{i}$>} are the coordinates
of the end point \emph{i}. All of them are \emph{Numeric} values.

\subsection{Region}

\begin{quotation}
(<face$_{1}$> <face$_{2}$> <face$_{n}$>)
\end{quotation}
where \emph{<face$_{i}$>} represents a face in the space, as a polygon
with (optionally) holes, and \emph{n}$\, \geq \, $0.

A face is represented as:

\begin{quotation}
(<outer\_cycle> <hole\_cycle$_{1}$> <hole\_cycle$_{2}$> $\cdots $
<hole\_cycle$_{n}$>)
\end{quotation}
where the \emph{outer\_cycle} and all the \emph{hole\_cycles} are
cycles (simple polygons) and \emph{n}$\, \geq \, $0 (this is, it
can be a face without holes).

A cycle is represented as:

\begin{quotation}
(<vertex$_{1}$> <vertex$_{2}$> $\cdots $ <vertex$_{n}$>)
\end{quotation}
being \emph{<vertex$_{i}$>} and \emph{<vertex$_{(i\, mod\, n)+1}$>}
consecutive vertices (and \emph{hence <vertex$_{1}$>}$\, \neq \, $\emph{<vertex$_{n}$>})
and \emph{n}$\, \geq \, $3.

A vertex is a \emph{Point} value.

\section{Representation of Time}
There are several representations for an \emph{instant} value.
All Applications using instants must be able to read each of this.
A \emph{DateTime} as an algebra type consist only of its value list. 
If \emph{DateTime} is used as subtype, e.g. in an interval, the
type-name has to be included. The typename can be {\tt datetime} or
{\tt instant}. Summarized we have three possibilities for a \emph{DateTime}:
\begin{tt}
  \begin{small}
     \begin{itemize}
       \item (datetime $<$datetimevalue$>$) \\
       \item (instant $<$datetimevalue$>$)  \\
       \item $<$datetimevalue$>$
      \end{itemize}
  \end{small}
\end{tt}

The representations of {\tt $<$datetimevalue$>$} are described
in the next subsections.

\subsection{String Representation}
\label{StringRepOfTime}
A $<$datetimevalue$>$ can be represented as string in the following
format: \\
\begin{tt}
  \begin{small}
      year-month-day[-hour:minute[:second[.millisecond]]]
  \end{small}
\end{tt}\\
The squared brackets indicate optionally parts. All contained values
are integers which must be in the appropriate interval.

\subsection{Real Representation}
A $<$datetimevalue$>$ can just be represented as an real value. 
The conversion beween the formats is described in section 
\ref{TimeConversions}.

\subsection{Julian Representation}
A $<$datetimevalue$>$ in the Julian representation is:\\
\begin{tt}
  \begin{small}
      (day millisecond)
  \end{small}
\end{tt}\\
{\tt day} can be any integer and millisecond is an integer in 
the interval [0,86400000[. {\tt day} represents the difference
to a fixed nulldate and {\tt milliseconds} represents the 
time part of this day.

\subsection{Gregorian Representation}
In the Gregorian Representation a $<$datetimevalue$>$ is written 
down as follow:\\
\begin{tt}
  \begin{small}
      (day month year [ hour minute [ second [ millisecond]]])
  \end{small}
\end{tt}\\
Like in the string representation (see section \ref{StringRepOfTime}),
the squared brackets are not part of the representation but indicates
optionally information. The values of the omitted parts are assumed to be zero.
The triple (day, month, year) must be a valid date. The remaining
values must be in the appropriate interval.

\subsection{Conversions between Representations of Time}
\label{TimeConversions}
The conversion between {\tt real} value and Julian representation
is straightforward:\\
$ JulToReal : int \times int \rightarrow \R$ \\
$ JulToReal(day,millisecond) = day + \frac{millisecond}{86400000}$ \\
The used operators are working on real numbers.\\[1em]
The another direction is computed as follow. Note a possible lost of 
precision.\\
$ RealToJul : \R \rightarrow (int,int)$ \\
$ RealToJul(instant) = (day,milli) $ \\
where:\\
{\small
$ day = \left\{ \begin{array}{ll}
                    int(instant) & \mbox{if } instant > 0 
                                   \mbox{ or } int(instant)=instant\\
                    int(instant)-1 & otherwise
                \end{array} \right. $\\[1em]
$ milli = \left\{ \begin{array}{ll}
                         tmp & \mbox{if } instant > 0 \mbox{ or } int(instant)=instant\\
                         tmp + 86400000 & \mbox{otherwise}
                      \end{array}\right. $ \\[1em]
\hspace*{1cm}where : tmp = int ((instant-int(instant))*86400000+0.5)}\vspace{1cm}

A quadrupel $(hour, min , sec, millisec)$ can be converted into
an single integer representing all milliseconds using the following formula:\\
$ millis : int \times int \times int \times int \rightarrow int $\\
$ millis(h,min,sec,msec) = (((h\cdot 60) + min)\cdot 60 + sec)\cdot 1000+ msec $
\\[1em]
The inverse direction can be computed as follow:\\
$ splitmillis: int \rightarrow (int,int,int,int)$ \\
$ splitmillis(ms) = (h,min,sec,msec)$ \\
where:
\begin{description}
   \item[] $ h =   {ms}/{3600000} \mod 24 $
   \item[] $ min = {ms}/{60000} \mod 60$
   \item[] $ sec = {ms}/{1000} \mod 60$
   \item[] $ msec = ms \mod 1000 $
\end{description}

The functions for converting a Julian Date into a Gregorian date and vice versa are
given in the appendix.

\section{Spatio-temporal Values}

Basically any spatio-temporal value of type \emph{<moving\_type>}
is represented as a string:

\begin{quotation}
(<unit$_{1}$> <unit$_{2}$> $\cdots $ <unit$_{n}$>)
\end{quotation}
where \emph{<unit$_{i}$>} is a unit value of the unit type corresponding
to \emph{<moving\_type>} and \emph{n}$\, \geq \, $0.

A unit value of a type \emph{<moving\_type>} is represented as:

\begin{quotation}
(<interval> <map\_type\_value>)
\end{quotation}

An \emph{<interval>} value is represented as:

\begin{quotation}
(<start> <end> <leftclosed> <rightclosed>)
\end{quotation}
where \emph{<start>} and \emph{<end>} are of type \emph{instant}
representing the start and end instant of the time interval, respectively, and
\emph{<leftclosed>} and \emph{<rightclosed>} are boolean values defining
if the interval is open (\emph{false}) or closed (\emph{true}) at
the start or end time instant, respectively.

\vspace*{3em}
The representation of a \emph{<map\_type\_value>} depends on the specific
\emph{type}.

\subsection{moving(point)}

The representation of a \emph{<map\_point\_value>} is:

\begin{quotation}
   (<x$_1$> <y$_1$> <x$_2$> <y$_2$>)
\end{quotation}
where x$_i$ and y$_i$ are \emph{Numeric} values. In a given Unit\\
\mbox{((<start> <end> <leftclosed> <rightclosed>) (<x$_1$> <y$_1$> <x$_2$> <y$_2$>))}
the point moves in the given interval from (x$_1$,y$_1$) to (x$_2$,y$_2$).
The position of a moving point can be computed for a single unit
using the following function:\\
 $ {pos} : {unit(point)} \times {instant} \rightarrow {point} $ \\
 $ {pos}(((s\ e\ l\ r) (x_1 y_1 x_2 y_2))),I) := \left\{
   \begin{array} {ll}
         \mbox{undefined} & \mbox{if } I \notin (s\: e\: l\: r) \\
         (x_1,y_1) & \mbox{if } s=e \mbox{\bf{} and } I \in (s\: e\: l\: r)\\
         (x_I,y_I) &  \mbox{otherwise}
   \end{array}
    \right. $\\[1em]

  where:\nopagebreak[4] \begin{description} \nopagebreak[4]
       \item $x_I = x_1+\frac{I-s}{e-s}\cdot(x_2 - x_1) $
       \item $y_I = y_1+\frac{I-s}{e-s}\cdot(y_2 - y_1) $
  \end{description}

\subsection{moving(points)}

The representation of a \emph{<map\_points\_value>} is:

\begin{quotation}
(<map\_point\_value$_{1}$> <map\_point\_value$_{2}$> $\cdots $
<map\_point\_value$_{n}$>)
\end{quotation}
with \emph{n}$\, >\, $0. If the \emph{movingpoints} value is empty
for a given time interval, no unit referring to such a time interval
appears in its representation.%
\footnote{This is the reason why \emph{n}$\, \neq \, $0.%
}

\subsection{moving(line)}

The representation of a \emph{<map\_line\_value>} is:

\begin{quotation}
(<map\_seg\_value$_{1}$> <map\_seg\_value$_{2}$> $\cdots $ <map\_seg\_value$_{n}$>)
\end{quotation}
where \emph{<map\_seg\_value$_{i}$>} is the mapping value of a segment
and n$\, >\, $0. If the \emph{movingline} value is empty for a given
time interval, no unit referring to such a time interval appears in
its representation.

A \emph{<map\_seg\_value>} is represented as:

\begin{quotation}
( \emph{<map\_point\_value$_{1}$>} \emph{<map\_point\_value$_{2}$>)}
\end{quotation}
representing the \emph{<map\_point\_value>} values of its end points.

\begin{description}
\item [Restriction:]a \emph{<map\_seg\_value>} must be defined over a plane
in the 3D space \emph{X}, \emph{Y}, \emph{Time}.
\end{description}

\subsection{moving(region)}

The representation of a \emph{<map\_region\_value>} is:

\begin{quotation}
(<map\_face\_value$_{1}$> <map\_face\_value$_{2}$> $\cdots $ <map\_face\_value$_{n}$>)
\end{quotation}
where n$\, >\, $0.

If the \emph{movingregion} value is empty for a given time interval,
no unit referring to such a time interval appears in its representation.

The representation of a \emph{<map\_face\_value>} is the same as the
spatial representation of a \emph{<face>}, but in this case each vertex
of its cycles is represented as a \emph{<map\_point\_value>}.

\begin{description}
\item [Restriction:]in a similar way as with the representation of a \emph{<map\_line\_value>},
any segment defined by two consecutive vertices \emph{<map\_point\_value$_{i}$>}
and \emph{<map\_point\_value$_{(i\, mod\, n)+1}$>} in a cycle of
a \emph{<map\_face\_value>} must be defined over a plane in the 3D
space \emph{X}, \emph{Y}, \emph{Time}.\end{description}

\appendix

\section{Conversion from a Gregorian Date to Julian Date and vice versa}

\begin{footnotesize}
\begin{verbatim}
#include <math.h>
static const long NULL_DAY = 2451547;
// this corresponds to 3.1.2004 and must be a multiple of 7

/*
   The function ToJulian computes the Julian day number of the given
   Gregorian date + the reference time.Positive year signifies A.D.,
   negative year B.C.Remember that the year after 1 B.C. was 1 A.D.
   Julian day 0 is a Monday.
   This algorithm is from Press et al., Numerical Recipes
   in C, 2nd ed., Cambridge University Press 1992
*/

static long ToJulian(int year, int month, int day){
  int jy = year;
  if (year < 0)
     jy++;
  int jm = month;
  if (month > 2)
     jm++;
  else{
     jy--;
     jm += 13;
  }

  int jul = (int)(floor(365.25 * jy) + floor(30.6001*jm)
                  + day + 1720995.0);
  int IGREG = 15 + 31*(10+12*1582);
  // Gregorian Calendar adopted Oct. 15, 1582
  if (day + 31 * (month + 12 * year) >= IGREG){
     // change over to Gregorian calendar
     int ja = (int)(0.01 * jy);
     jul += 2 - ja + (int)(0.25 * ja);
  }
  return jul-NULL_DAY;
}

/*
   This function converts a Julian day to a date in the Gregorian calender.

   This algorithm is from Press et al., Numerical Recipes
   in C, 2nd ed., Cambridge University Press 1992

*/
static void ToGregorian(long Julian, int &year, int &month, int &day){
   int j=(int)(Julian+NULL_DAY);
   int ja = j;
   int JGREG = 2299161;
   /* the Julian date of the adoption of the Gregorian
      calendar
   */
    if (j >= JGREG){
    /* cross-over to Gregorian Calendar produces this
       correction
    */
       int jalpha = (int)(((float)(j - 1867216) - 0.25)/36524.25);
       ja += 1 + jalpha - (int)(0.25 * jalpha);
    }
    int jb = ja + 1524;
    int jc = (int)(6680.0 + ((float)(jb-2439870) - 122.1)/365.25);
    int jd = (int)(365 * jc + (0.25 * jc));
    int je = (int)((jb - jd)/30.6001);
    day = jb - jd - (int)(30.6001 * je);
    month = je - 1;
    if (month > 12) month -= 12;
    year = jc - 4715;
    if (month > 2) --year;
    if (year <= 0) --year;
}

\end{verbatim}
\end{footnotesize}

\end{document}