### 2006-02-22 Algebras/MovingRegion/Intro/Intro.tex
Added Intro document.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

\documentclass[a4paper,12pt]{article}

\usepackage{graphicx}
\usepackage{caption2}

\title{Using the {\tt MovingRegionAlgebra}}
\author{Holger M\"{u}nx {\tt $<$hm@muenx.net$>$}
    \thanks{The {\tt MovingRegionAlgebra} has been developed
    as bachelor thesis with Prof.\ Dr.\ Ralf Hartmut G\"{u}ting,
    Fernuniversit\"{a}t Hagen, Fachbereich Informatik, 
    Lehrgebiet Praktische Informatik IV}}
\date{22-Feb-2005 / INCOMPLETE DRAFT}

\newcommand{\secondo}{{\scshape SE\-CON\-DO}}
\newcommand{\mra}{{\tt MovingRegionAlgebra}}
\newcommand{\mr}{{\tt movingregion}}

\setlength{\parindent}{0em}
\setlength{\parskip}{1.5ex}

\renewcommand{\captionfont}{\bf}

\begin{document}

\maketitle

\begin{abstract}
This document provides a number of practical examples for the usage
of the \mra\ in the \secondo\ DBMS. It describes
both database objects, related queries and the application of
\secondo's Hoese viewer to display objects and query results.
\end{abstract}

\section{Introduction}

Spatial and temporal objects in databases are used to represent
real physical objects, which have a position and extension in space
as attributes, which can change over time. Examples of such objects
are countries, mountains and rivers or planes, cars and ships.
Pollution distribution or storm areas are examples for moving
regions. \cite{GuS04}, \cite{EGS+99}, \cite{GBE+00}, \cite{FGN+00}
and \cite{LFG03}
provide introductions to temporal objects on various levels.

\secondo\ is an extensible DBMS and is introduced in \cite{DiG99}.
Algebras can be added to \secondo\ to implement additional data types
and operators. The {\tt SpatialAlgebra}
and the \mra\ support spatial and selected temporal
objects. The \mra\ implements moving regions and
is described in \cite{Mue05}.

This document provides information about using the \mra. The
focus is on a very practical perspective, including numerous examples
and screenshots. Basic understanding of \secondo's operation is required
to apply this document.
Section \ref{creating} explains methods to create
\mra\ objects. Section \ref{operations} describes operations on
\mra\ objects including different ways to visualise the results.

\section{Creating {\tt movingregion} objects}
\label{creating}

\subsection{Simple objects}

The following \secondo\ statement creates a simple \mr\ object {\tt mr1},
consisting
of a single region unit, which contains one face with a cycle and a hole
cycle. The unit's interval ranges from instant $0.0$ to instant $10.0$,
which are the
floating point representations of the instants {\tt 2000-01-03-00:00:00.000}
and {\tt 2000-01-03-00:00:00.000}.
\begin{verbatim}
    let mr1 = 
      [ const movingregion value 
          (((0.0 10.0 TRUE TRUE)
            ((((1.0 3.5 1.5 1.5)
               (2.0 5.5 3.0 4.5)
               (3.0 6.5 3.5 5.0)
               (4.0 6.5 5.5 5.0)
               (4.0 5.5 5.5 4.5)
               (5.0 4.5 7.5 2.5)
               (5.0 2.5 7.5 1.0)
               (4.0 1.5 7.0 0.5)
               (3.0 1.5 2.5 0.5))
              ((2.0 3.0 3.0 2.0)
               (2.0 4.0 3.0 3.0)
               (3.0 4.0 4.0 3.0)
               (3.0 3.0 4.0 2.0)))))) ];
\end{verbatim}
Figure
\ref{creating:simple1} shows the object in the Hoese viewer in the initial
instant. A complete description of the list representation of \mr\ objects is
presented in \cite{LeB04}.
\begin{figure}
  \begin{center}
    \includegraphics[scale=0.4]{screenshot01.png}
  \end{center}
  \caption{A simple \mr\ object in the Hoese viewer}
  \label{creating:simple1}
\end{figure}

Region units may degenerate in two different ways:
\begin{itemize}
  \item Segments may be reduced
    to points in its initial or final instant (but not in both).
  \item A segment may be equal to other segments in
    the initial or final instant. A segment may be even equal to 
    segments in the initial and the final instant as long as it is not
    equal to the same segment in the initial and the final instant.
\end{itemize}
A region unit must not degenerate in other instants than its initial
or final instant. Section 4.3.3 in \cite{GuS04} and other literature
requires to use open
intervals for the degenerated instants; the current implementation of
the \mra\ is more flexible and treats degeneration properly.
Figure \ref{creating:simple2} contains examples of
degeneration in three region units.
\begin{figure}
  \begin{center}
    \includegraphics[scale=0.75]{uregion-degenerated.pdf}
  \end{center}
  \caption{Degenerated region units}
  \label{creating:simple2}
\end{figure}

The following statement creates a \mr\ object, whose single region
unit resembles the rightmost region unit in figure \ref{creating:simple2}.
\begin{verbatim}
    let mr5 =
      [ const movingregion value
          (((0.0 10.0 TRUE TRUE)
            ((((2.0 1.0 1.0 4.0)
               (2.0 3.0 1.0 7.0)
               (2.0 3.0 3.0 6.0)
               (6.0 3.0 5.0 6.0)
               (6.0 3.0 7.0 6.0)
               (6.0 1.0 7.0 4.0)
               (6.0 1.0 5.0 6.0)
               (2.0 1.0 3.0 6.0)))))) ];
\end{verbatim}
In the initial instant, this object should be a region, which has the
shape of a rectangle. Lets apply the {\tt initial} operator to verify this.
\begin{verbatim}
    Secondo => query initial(mr5);

    No specific display function defined. Generic function used.
    Type: intimeregion
    Value: 
    ("2000-01-03" 
        (
            (
                (
                    (6.0 1.0) 
                    (2.0 1.0) 
                    (2.0 3.0) 
                    (6.0 3.0)))))
\end{verbatim}
This is exactly the expected result. Applying the {\tt final} operator
results in two faces, which have the shape of two triangles pointing to
each other, which is again exactly the expected result.
\begin{verbatim}
    Secondo => query final(mr5);

    No specific display function defined. Generic function used.
    Type: intimeregion
    Value:
    ("2000-01-13" 
        (
            (
                (
                    (3.0 6.0) 
                    (1.0 4.0) 
                    (1.0 7.0))) 
            (
                (
                    (7.0 4.0) 
                    (5.0 6.0) 
                    (7.0 6.0)))))
\end{verbatim}

\subsection{Object creation in the Hoese viewer}

\subsection{The {\tt Region2MRegion} converter}

\section{Operations on {\tt movingregion} Objects}
\label{operations}

\begin{thebibliography}{\hspace{2cm}}

\bibitem[DiG99]{DiG99}
  S.\ Dieker, R.\ H.\ G\"{u}ting.
  \textsl{Plug and Play with Query Algebras: \secondo\ A Generic DBMS
    Development Environment}.
  Informatik Berichte 249 - 2/1999, Fernuniversit\"{a}t Hagen, 1999.

\bibitem[EGS+99]{EGS+99}
  M.\ Erwig, R.\ H.\ G\"{u}ting, M.\ Schneider, M.\ Vazirgiannis.
  \textsl{Spatio-Temporal Data Types: An Approach to Modeling and
    Querying Moving Objects in Databases}.
  GeoInformatica 3:3, 1999.

\bibitem[FGN+00]{FGN+00}
  L.\ Forlizzi, R.\ H.\ G\"{u}ting, E.\ Nardelli, M.\ Schneider.
  \textsl{A Data Model and Data Structures for Moving Objects Databases}.
  Proceedings of the 2000 ACM SIGMOD International Conference on Management
  of Data, 2000.

\bibitem[GBE+00]{GBE+00}
  R.\ H.\ G\"{u}ting, M.\ H.\ B\"{o}hlen, M.\ Erwig, C.\ S.\ Jensen,
  N.\ A.\ Lorentzos, M.\ Schneider, M.\ Vazirgiannis.
  \textsl{A Foundation for Representing and Querying Moving Objects}.
  ACM Transactions on Database Systems, Vol. 25, No. 1, March 2000.

\bibitem[GuS04]{GuS04}
  R.\ H.\ G"uting, M.\ Schneider.
  \textsl{Moving Objects Databases}.
  Vorlesung an der Fernuniversit"at Hagen, 2004.

\bibitem[LeB04]{LeB04}
  J.\ A.\ C.\ Lema, T.\ Behr.
  \textsl{External Representation of Spatial and Spatio-Temporal Values}.

\bibitem[LFG03]{LFG03}
  J.\ A.\ C.\ Lema, L.\ Forlizzi, R.\ H.\ G\"{u}ting, E.\ Nardelli,
  M.\ Schneider.
  \textsl{Algorithms for Moving Objects Databases}.
  The Computer Journal, Vol. 46, No. 6, 2003.

\bibitem[Mue05]{Mue05}
  H.\ M\"{u}nx.
  \textsl{Behandlung zeitabh\"{a}ngiger Gebiete in Datenbanken
    f\"{u}r bewegte Objekte}.
  Bachelor thesis, Fernuniversit\"{a}t Hagen, 2005.

\end{thebibliography}

\end{document}