### 2006-03-02 Algebras/Periodic/Figures/SecondoLogo_command.tex
some files dor commenting the periodic algebra

% This file provides the logo of the
% Secondo system 
%
% for using this command, the packages
% pstricks,and  relsize has to be imported
%
% This command understands a single parameter defining the
% width of this logo.
%
% Note that the unit sizes ofthe pstricks packages will be changed
% by using this command.

\newcommand{\SecondoLogo}[1]{
% define the colors of the picture
\newrgbcolor{shadow}{0.8046875 0.79296875 0.8125}
\newrgbcolor{secred}{0.7421875 0 0}
\newrgbcolor{purple}{0.9453125 0.796875 0.796875}
\newrgbcolor{secblack}{0 0 0}
\newdimen\logowidth
\logowidth=#1
\psset{unit=1mm}
\scaleboxto(\logowidth,0){
\begin{pspicture}(0,0)(118, 21) % (1430 256)
\rput[bl]{0}(0,2){
% First : paint all shadows
% the text
\rput[bl]{0}(0.7,-0.3){\scaleboxto(117,17){\textsf{\shadow{\larger S}{\smaller ECONDO}}}}
% rectangle below
\rput[bl]{0}(1,-1){\pspolygon[fillstyle=solid,fillcolor=shadow,linestyle=none](0,0)(117,0)(117,2.5)(0,2.5)}
% rectangle above
\rput[bl]{0}(1,-3.5){\pspolygon[fillstyle=solid,fillcolor=shadow,linestyle=none](17,18)(117,18)(117,20)(17,20)}

% the black parts
% rectangle below
\rput[bl]{0}(0,-0.5){\pspolygon[fillstyle=solid,fillcolor=black,linestyle=none](0,0)(117,0)(117,2.5)(0,2.5)}
% rectangle above
\rput[bl]{0}(0,-3){\pspolygon[fillstyle=solid,fillcolor=black,linestyle=none](17,18)(117,18)(117,20)(17,20)}
% the text
\rput[bl]{0}(0,0){\scaleboxto(117,17){\textsf{\secred{\larger S}{\smaller ECONDO}}}}

% the inverse text
\begin{psclip}{
  \rput[bl]{0}(0,-0.5){\pspolygon[linewidth=0.3pt](0,0)(117,0)(117,2.5)(0,2.5)}}
  {\rput[bl]{0}(0,0){\scaleboxto(117,17){\textsf{\purple{\larger S}{\smaller ECONDO}}}}}
\end{psclip}
% circles and lines in the rectangle above
\rput[l]{0}(0,0){
	% dots above the 'E'
	\rput[l](19,16){\pscircle[fillstyle=solid,linestyle=none,fillcolor=purple](0,0){0.7}}
	\rput[l](23.5,16){\pscircle[fillstyle=solid,linestyle=none,fillcolor=purple](0,0){0.7}}
	\rput[l](28,16){\pscircle[fillstyle=solid,linestyle=none,fillcolor=purple](0,0){0.7}}
	% rectangles above CON
	\rput[l](33,16){\pspolygon[fillstyle=solid,fillcolor=purple,linestyle=none](0,-0.35)(11,-0.35)(11,0.35)(0,0.35)}
	\rput[l](49,16){\pspolygon[fillstyle=solid,fillcolor=purple,linestyle=none](0,-0.35)(11,-0.35)(11,0.35)(0,0.35)}

	\rput[l](66,16){\pspolygon[fillstyle=solid,fillcolor=purple,linestyle=none](0,-0.35)(11,-0.35)(11,0.35)(0,0.35)}
	% dots above the 'D'
	\rput[l](84,16){\pscircle[fillstyle=solid,linestyle=none,fillcolor=purple](0,0){0.7}}
	\rput[l](88.5,16){\pscircle[fillstyle=solid,linestyle=none,fillcolor=purple](0,0){0.7}}
	\rput[l](93,16){\pscircle[fillstyle=solid,linestyle=none,fillcolor=purple](0,0){0.7}}
	% rectangle above 'O'
	\rput[l](101,16){\pspolygon[fillstyle=solid,fillcolor=purple,linestyle=none](0,-0.35)(11,-0.35)(11,0.35)(0,0.35)}
}
} % rput
\end{pspicture}}
}