### 2008-02-02 Algebras/RegionInterpolation/Tex/Quelltexte.tex

\chapter{Quelltexte}
\minitoc
\newpage
\begin{landscape}
\section{ConvexHullTreeNode}
\begin{tiny}
\lstinputlisting[multicols=2,numbers=none]{ConvexHullTreeNode.java}
\end{tiny}
\end{landscape}