### 2008-02-26 Algebras/RegionInterpolation/Tex/Quelltexte.tex

\chapter{Quelltexte}
\minitoc
\newpage
\begin{landscape}
\section{ConvexHullTreeNode}
\begin{tiny}
\lstinputlisting[language=c++,multicols=2,numbers=none]{Lines15.cpp}
\end{tiny}
\end{landscape}