### 2009-06-15 Optimizer/Subqueries/kapitel/einleitung.tex
sketches of thesis structure

%
% Einleitung
%
%

\chapter{Einleitung}

\pagenumbering{arabic}

\section{Beschreibung SECONDO}
\begin{itemize}
	\item Erweiterbares Datenbanksystem
	\item Second-Order-Signatures
	\item Kernel/Algebren
	\item Optimierer
	\item GUI
\end{itemize}

\section{SQL}
\begin{itemize}
	\item Datenbanksprache
	\subitem Datendefinition (DDL)
	\subitem Datenmanipulation (DML)
	\subitem Rechtevergabe (DCL)
	\item Abfrageblock bestehend aus 
	\subitem SELECT-Klausel
	\subitem FROM-Klausel
	\subitem WHERE-Klausel
	\subitem GROUP BY\dots HAVING-Klausel
	\subitem ORDER BY-Klausel
	\item geschachtelte Abfragen
	\subitem in der SELECT-Klausel
	\subitem in der FROM-Klausel
	\subitem in der WHERE Klausel
	\subitem in der GROUP BY\dots HAVING-Klausel
\end{itemize}

%
% EOF
%
%