### 2009-07-26 Optimizer/Subqueries/kapitel/einleitung.tex
Current version

%
% Einleitung
%
%

\chapter{Einleitung}
\pagenumbering{arabic}
\section{Problembeschreibung}

%
% EOF
%
%