### 2009-08-23 Optimizer/Subqueries/kapitel/einleitung.tex
daily update

%
% Einleitung
%
%

\chapter{Einleitung}\label{chp:Einleitung}
\pagenumbering{arabic}
\section{Problembeschreibung}\label{sct:Problembeschreibung}
\section{Aufbau der Arbeit}\label{sct:Aufbau der Arbeit}

%
% EOF
%
%