### 2013-10-10 CM-Scripts/studentuserlist.tex
allowed file size temporarily increased

\documentclass[a4paper,9pt,landscape]{scrartcl}
\usepackage[top=1cm,bottom=1cm,left=2cm,right=1cm]{geometry}
\usepackage[utf8x]{inputenc}
\usepackage[ngerman]{babel}
\usepackage{tabularx}
\usepackage{tabu}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{color}
\usepackage{colortbl}
\begin{document}
\section{CVS-Zug\"ange f\"ur Studenten, Stand: \today}
\begin{longtabu}to 1\textwidth{|l|l|l|l|l|X|l|l|}
  \hline \hline
\textbf{Linux-User}        &\textbf{CVS-User}        &\textbf{Student}&\textbf{E-Mail}       &\textbf{Verzeichnis}              &\textbf{von} &\textbf{bis}\\
\hline
\hline\endhead
\rowcolor{yellow}\cellcolor{red}diplomand1   &alber  &Thomas Alber    &th.alber@loop.de      &Javagui/viewer/queryconstruction2 &17.06.13     &\\
\cline{2-7}
\cellcolor{red}            &rentergent               &Lisa Rentergent &lisa@rentergent.de    &Javagui/viewer/queryconstruction  &16.04.12     &17.06.13\\
\cline{2-7}
\cellcolor{red}            &uchdorf                  &Thomas Uchdorf  &                      &Algebras/OSM                      &08.02.11     &05.03.12\\
\cline{2-7}
\cellcolor{red}            &vjohn                    &Volker John     &                      &Algebras/OptAux                   &             &08.02.11\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand2   &ahmed  &Jasmine Ahmed   &j.ahmed@gmx.de        &Javagui/viewer/update2            &19.06.13     &\\
\cline{2-7}
\cellcolor{red}            &keberl                   &Katharina Rieder&keberl@web.de         &Algebras/TrajetoryAnnotation      &16.04.12     &19.06.13\\
\cline{2-7}
\cellcolor{red}            &fpanse                   &Frank Panse     &                      &Algebras/SymbolicTrajectory       &21.12.10     &05.03.12\\
\cline{2-7}
\cellcolor{red}            &nvoget                   &Nicolai Voget   &                      &Algebras/OrderedRelation          &             &21.12.10\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand3   &eckebre&Jens Eckebrecht &jens@eckebrecht.de    &Algebras/RobustPlaneSweep         &21.12.12     & \\
\cline{2-7}
\cellcolor{red}            &kubusch                  &Rico Kubusch    &rico.kubusch@bka.bund.de&                                &21.06.11     &21.12.12\\
\cline{2-7}
\cellcolor{red}            &bponeleit                &Burkhard Poneleit&                     &Optimizer/Subqueries              &             &08.02.11\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand4   &walther&Daniel Walther  &da.walther@gmx.de     &Tools/SecondoPostGIS              &22.02.13     &\\
\cline{2-7}
\cellcolor{red}            &kkoch                    &Katja Koch      &katjakoch1@googlemail.com&Algebras/RobustGeometry        &02.01.12     &21.12.12\\
\cline{2-7}
\cellcolor{red}            &geduldig                 &Jochen Geduldig &                      &Algebras/NestedRelation2          &08.09.10     &02.01.12\\
\cline{2-7}
\cellcolor{red}            &kteufel                  &Klaus Teufel    &                      &Algebras/NestedRelation           &             &08.09.10\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand5   &zacher &Dirk Zacher     &dirk.zacher@fernuni-hagen.de&Algebras/Tile \& Algebras/Raster2&18.04.13     &\cellcolor{red}Laptop\\
\cline{2-7}
\cellcolor{red}            &tachmann                 &Thomas Achmann  &t\_achmann@gmx.de     &Algebras/Distributed              &16.02.11     &21.12.12\\
\cline{2-7}
\cellcolor{red}            &totim                    &Tobias Timmerscheidt&                  &Algebras/DistributedRelation      &15.01.10     &16.02.11\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand6   &floheinz&Florian Heinz  &fh@sysv.de            &Algebras/RegionInterpolation2 \& Javagui/viewer/v3d &26.04.13 &\\
\cline{2-7}
\cellcolor{red}            &srenner                  &Stefanie Renner &stefanie.renner@gmx.de&Algebras/MovingRegion2            &09.11.11     &21.12.12\\
\cline{2-7}
\cellcolor{red}            &sjung                    &Sven Jungnickel &                      &Algebras/ExtRelation2             &             &08.02.11\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand7   &steigerk&Kristina Steiger&tinasteiger@gmx.net  &WebGui                            &06.05.13     &\\
\cline{2-7}
\cellcolor{red}            &geisbues                 &Simon Geisbuesch&simon.geisbuesch@o2online.de&Algebras/SpatialJoin2       &14.11.11     &21.12.12\\
\cline{2-7}
\cellcolor{red}            &zdahmen                  &Zoryana Dahmen-Lobur&                  &/DahmenLobur                      &20.10.10     &08.02.11\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand8   &kuepper&Michael K\"upper&kuepper.michael@gmail.com&AndroidGui                     &06.05.13     &\\
\cline{2-7}
\cellcolor{red}            &dcapek                   &Dieter Capek    &d007@capek.de         &Algebras/Groupby                  &16.11.11     &21.12.12\\
\cline{2-7}
\cellcolor{red}            &linsenme                 &Linsenmeier     &                      &Algebras/SpatialJoin \& Algebras/NearestNeighbour& &08.02.11\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand9   &issa   &Hamza Adnan Issa&hamza.adnan.issa@gmail.com&Algebras/HybridTrajectory     &11.06.13     &\\
\cline{2-7}
\cellcolor{red}            &mroth                    &Matthias Roth   &fernstudium@planet-roth.de&Algebras/MapMatching          &21.11.11     &21.12.12\\
\cline{2-7}
\cellcolor{red}            &whoffmann                &Walter Hoffmann &                      &secondo\_Whoff                    &01.02.10     &08.02.11\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand10  &dbecker&Dirk Becker     &tyr@tyrbedarf.de      &Algebras/Cluster2                 &19.06.13     &\\
\cline{2-7}
\cellcolor{red}            &willmes                  &Gero Willmes    &student@harpspieler.de&Optimizer/LargeQueries            &29.11.11     &21.12.12\\
\cline{2-7}
\cellcolor{red}            &herkert                  &A. Herkert      &                      &LargeQueries                      &01.03.10     &08.02.11\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand11  &schroeer&Susanne Schröer&susanne@team-schroeer.com&(noch keins)                   &19.08.13     &\\
\cline{2-7}
\cellcolor{red}            &schrage                  &Michael Schrage &schragman@googlemail.com&JDBC                            &23.12.11     &21.12.12\\
\cline{2-7}
\cellcolor{red}            &danbro                   &Daniel Brockmann&                      &Algebras/UGrid                    &14.04.10     &08.02.11\\
\hline
\hline
\cellcolor{green}diplomand12&mtirtasana              &Matthias Tirtasana&matthias.tirtasana@science-computing.de&OptParser      &05.03.12     &21.12.12\\
\cline{2-7}
\cellcolor{green}          &wrede                    &Nina Wrede      &                      &Algebras/Network2 \& Algebras/TemporalNet2&15.06.10&02.01.12\\
\hline
\hline
\rowcolor{yellow}\cellcolor{green}diplomand13&nidzwetzki&Jan Nidzwetzki&knidzwetzki@gmx.de   &Algebras/CostEstimation           &18.04.12     &\\
\hline
\hline
\cellcolor{green}diplomand14&flehmann                &Frank Lehmann   &fraleco@gmx.de        &Algebras/SymbolicTrajectory2      &23.04.12     &12.12.12\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand15&nvkempen &Nicolai van Kempen&nikolaivank@googlemail.com&Optimizer/NestedRelations \& Optimizer/MemoryAllocation&24.04.12&\\
\hline
\hline
\cellcolor{green}diplomand16&jmoeller                &Jasmin Moeller  &jasminmoeller@gmx.de  &Algebras/Cluster2                 &28.06.12     &21.12.12\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand17&ofeuer   &Oliver Feuer    &ofeuer\_informatik@yahoo.de&Algebras/Region2 \& MovingRegion3&11.10.12 &\cellcolor{red}Laptop\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand18&sfuhs    &Svenja Fuhs     &svenja.fuhs@t-online.de&Algebras/Precise2D               &29.10.12     &\\
\hline
\hline
\rowcolor{yellow}\cellcolor{red}diplomand19&ehnes    &J\"urgen Ehnes  &juergen@ehnes.eu      &                                  &20.11.12&\\
\hline
\hline
\cellcolor{green}diplomand20&                        &                &                      &                                  &             &\\
\hline
\hline
\end{longtabu}
\end{document}